---
title: "Lab week 8"
author: "Teresa Fukuda"
date: "11/22/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

#load the tidyverse and read in the salmon_imports data

library(tidyverse)

salmon <- read_csv("salmon_imports.csv")

```

####Exploratory scatterplot
 
```{r}

salmon_scatter <- ggplot(salmon, aes(x=year, y=salmon_fresh)) +
  geom_point()
salmon_scatter

# Overall, this data has a very linear trend except for a weird deviation in about 2004 and 2012; generally it makes sense that the fresh salmon would increase linearly, probably only until a certain point where salmon stocks are limiting.

```
 
####Attempting a linear regression

```{r}

salmon_model <- lm(salmon_fresh ~ year, data=salmon)

salmon_model

# model linear formula: salmon imports (thousands) = -14982940 + 7550(year)


```

The linear model give an equation with an intercept of -14982940 and a slope of 7550. That means that every year, based on the data we have, the model predicts that the salmon imports will increase by 7550 thousand pounds. The y-intercept is super negative, meaning that if we extrapolate far into the past, the model says that we were adding huge amounts of salmon into the ocean. 


####Model diagnostics
```{r}
plot(salmon_model)

# The residuals appear to be normally distributed, based on the residual vs fitted plot that does not appear to have a patterns and the normal q-q that is very linear. 

```

####Model fit and significance

```{r}

summary(salmon_model)


```

Looking at the summary of the model, with an r-squared value of 0.9478, almost 95% of the variation in the salmon imports can be predicted by the year. With a p value < 0.001, I can say that the year significantly predicts salmon imports. There is 5.22% of variation in the salmon imports that cannot be explained by the year. None of this implies any kind of causation! Careful with wording here.

#### Prepare a final graph. 

```{r}

salmon_final_graph <- ggplot(salmon, aes(x=year, y=salmon_fresh)) +
  geom_point() +
  geom_smooth(method=lm, se=TRUE, size=0.5, color = "gray20")+
  theme_bw() +
  scale_x_continuous(expand = c(0,0), limits = c(1990, 2017)) +
  labs(x="Year", y= "U.S. Annual Fresh Salmon Imports \n(1,000 pounds)")

salmon_final_graph

```


#### Making predictions using the model

```{r}

# create new dataframe with column "years" containing the ones that we want to predict
new_years<-data.frame(year=c(2022,2024,2026,2028))

# use this dataframe and the predict() function to find the new values for the model

future_predict <- predict(salmon_model,newdata=new_years,interval= "confidence")
future_predict

#bind this data and "new_years" data for a useful table of data

predictions <- data.frame(new_years, future_predict)
predictions

```

